Click chemistry reveals an alternate mechanism for HPII catalase

### 2018-02-22 bb494/manuscript.tex
finish intro?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Catalase is an enzyme which catalyzes the reaction $2H_2O_2 \rightarrow O_2 + 2H_2O$. This is an important detoxification reaction, since hydrogen peroxide is a reactive toxin generated by several biological processes. Catalase is a tetrameric protein which uses a heme group in its active site to facilitate the breakdown of hydrogen peroxide. It is a unique protein in that its reaction evolves gas, and thus its activity can be seen with the bare eye. The protein has been studied as a target for tumor suppression drugs \cite{bauer}. The enzyme is also useful industrially, in particular in the bleaching of textiles \cite{amorim}.

The molecular mechaninism of catalase proceeds in two steps. First, the heme reduces a hydrogen peroxide, producing water and the so-called Cpd I. Then Cpd I oxidizes a second hydrogen peroxide, producing water and oxygen gas. This second step is mediated by a histidine near the active site \cite{alfonso-prieto}.

@@

Catalase is an enzyme which catalyzes the breakdown of hydrogen peroxide into water and oxygen gas. This is an important detoxification reaction, since hydrogen peroxide is a reactive toxin generated by several biological processes. Catalase is a tetrameric protein which uses a heme group in its active site to facilitate the breakdown of hydrogen peroxide. The protein is

Catalase's activity is nearly diffusion-limited \cite{kcatkm, difflimit}. This rate suggests that nearly every step in the substrate-enzyme interaction is optimized for speed. Understanding these optimizations is good because __________. Catalase has a variety of interesting optimizations in funneling substrate into its active pocket. The enzyme has a number of high-H2O2-residency surface residues near its active channel, which are believed to concentrate the substrate near the entrance to the active channel \cite{concentrateh2o2}. The active channel itself seems optimized to maximize substrate flow, to the point that mutagenesis to open the channel decreases enzyme activity \cite{substrateflow}. The channel seems to be capable of selecting the molecules which enter it using a patch of hydrophobic residues which are situated to preferentially pass the bulkier H2O2 \cite{molecularruler}. Even the directionality of substrate within the channel seems to be controlled: one channel is used for substrate entrance, and the other for product exit \cite{lateralchannel}. An new finding is that catalase might control not just the presence of catalase, but even its orientation as it enters the active site. It is believed that an electric field from a negative Asn residue to the positive heme acts to orient the H2O2 dipole, positioning the molecule such that it enters the active site at an angle ideal for reacting \cite{electricpotential}. It is this orientation that we hope to explore in this research.

The study on the electric potential in the active channel \cite{electricpotential} hypothesized the existance of an orientation potential, but ___flaws_in_experiment__. We hope to more thoroughly probe this effect using genetic code expansion ____ need to figure out how this mutation should influence the active site.

@@

@@

An Electrical Potential in the Access Channel of Catalases Enhances Catalysis, loewen
An electrical gradient between Asp181 and the positive heme orients the waters in the pocket, promoting H-bonding and rotating H2O2 properly

Want: 6-8 substrate concentrations; 3 repeats

Hypothesis: can we control catalase catalytic efficiency by modulating dielectric constant, thus dampening the effects of electric field between Asp and the heme?
salt vs dielectric const: Dielectric Relaxation of Aqueous NaCl Solutions, Buchner, May, 10.1021/jp982977k
temp vs dielectric const: Dielectric Constant of Water from 0 0 to 1000 C , Malmberg, Mariott, eps = 87. 740 - 0.400081t + 9.398(10^-4)t^2 + 1.410(10^-6)t^3

Diffusion-Controlled Reactions of Enzymes, Wei-Zhu Zhong
klim: 1.8*10^-9

EFFECT OF pH UPON THE REACTION KINETICS OF THE ENZYME-SUBSTRATE COMPOUNDS OF CATALASE*
Catalase reaction rate is constant at pH 5-9

New hypothesis: Catalase supposedly benefits from an electric field pointing through its active channel to properly position its substrate. We hypothesize that by destroying this electric field by inserting negatively charged groups near the heme, we can lower the reaction rate, possibly below the diffusion limit.
Catalase has constant activity between pH 5 through 9. We will insert a group with pKa of 7, which should be neutral at pH 6 and mostly positive at pH 8. If such an enzyme experiences a drop in catalytic efficiency between pH 7 and 9, this will be evidence of the importance of an electric field orienting substrate molecules.

3NY pka: 7.230 on the hydroxy group, Conversion of 3-nitrotyrosine to 3-aminotyrosine in peptides and proteins, Vallee

@@

The Catalase-Hydrogen Peroxide System, Suggett
Km: 1.1M,
kcat/Km: 6.6*10^7/M/s

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Catalase is an enzyme which catalyzes the reaction $2H_2O_2 \rightarrow O_2 + 2H_2O$. This is an important detoxification reaction, since hydrogen peroxide is a reactive toxin generated by several biological processes.

@@

Catalase is an enzyme which catalyzes the breakdown of hydrogen peroxide into water and oxygen gas. This is an important detoxification reaction, since hydrogen peroxide is a reactive toxin generated by several biological processes. Catalase is a tetrameric protein which uses a heme group in its active site to facilitate the breakdown of hydrogen peroxide. The molecular mechaninism of catalase proceeds in two steps. First, the heme reduces a hydrogen peroxide, producing water and the so-called Cpd I. Then Cpd I oxidizes a second hydrogen peroxide, producing water and oxygen gas. This second step is mediated by a histidine near the active site \cite{alfonso-prieto}.

Catalase's activity is nearly diffusion-limited \cite{kcatkm, difflimit}. This rate suggests that nearly every step in the substrate-enzyme interaction is optimized for speed. Understanding the optimizations of diffusion-limited proteins can provide many insights into how proteins achieve their functions. Catalase has a variety of interesting quirks for funneling substrate into its active pocket. The enzyme has a number of high-H2O2-residency surface residues near its active channel, which are believed to concentrate the substrate near the entrance to the active channel \cite{concentrateh2o2}. The active channel itself seems optimized to maximize substrate flow, to the point that mutagenesis to open the channel decreases enzyme activity \cite{substrateflow}. The channel seems to be capable of selecting the molecules which enter it using a patch of hydrophobic residues which are situated to preferentially pass the bulkier H2O2 \cite{molecularruler}. Even the directionality of substrate within the channel seems to be controlled: one channel is used for substrate entrance, and the other for product exit \cite{lateralchannel}. An new finding suggests that catalase might control not just the presence of hydrogen peroxide, but even its orientation as it enters the active site. It is believed that an electric field from a negative Asn residue to the positive heme acts to orient the H2O2 dipole, positioning the molecule such that it enters the active site at an angle ideal for reacting \cite{electricpotential}. This use of an electric field over a long range to create an ``orienting field'' is a new and exciting usage of residues to achieve a function; understanding its role in catalase may help to understand the mechanisms of other proteins.

In this study we hope to study the proposed orienting field optimization and its influence on catalase activity. In particular, we hope to carefully disrupt the wildtype active site potential environment by introducing a new charged residue, then examine the effect this disruption has on activity. For meaningful results, we must be careful to not introduce structural modifications to the enzyme. To ensure this, we will use genetic code expansion to incorporate 3-nitrotyrosine in place of tyrosine 206. The T206NY mutation will add only 30 $\angstrom^3$ of steric bulk \cite{3ntsize}, so should have minimal influence on active site structure. The theorized results of this addition are diagrammed in Figure \ref{fig:hypothesis}. To study the influence of field disruption on activity, we will study the activity of T206NY above and below its pKa of 6.8 \cite{3ntsize}. Wildtype catalase activity is constant under pH values from 5 to 10 \cite{phdependence,kcatkm}, so pH-dependence should only be seen if electric environment in the active site is important for activity.\\

\begin{figure}
  \caption{Depiction of catalase active pocket. \textbf{Left:} wildtype pocket with proposed ligand-orienting E-field \cite{electricpotential}. The field should orient H2O2 for optimal reaction with the heme. \textbf{Right:} modified HPII-3NY with a deprotonated 3-nitrotyrosine residue at site 206, with hypothetical augmented E-field. The disrupted E-field should result in un- or mis-aligned substrate.}
  \label{fig:hypothesis}
\end{figure}

@@

\textbf{}

@@

@@

\cite{kcatkm}
Km: 1.1M,
kcat/Km: 6.6*10^7/M/s

possibly make a figure with the h2o2 and water from the electric potential paper?

---------------
Arguments about protein production, validation, and ncAA validation:
We know we have wildtype protein because our gels have bands, indicating protein of the proper size was made with His tags
We know our WT protein was active by quantitative assays

We know our noncanonical protein was expressed by similar
